---
title: "Math Lessons"
author: "R-Girls"
output: 
   learnr::tutorial:
    css: css/custom.css
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(tibble)
library(fontawesome)
library(astsa)
library(grid)
knitr::opts_chunk$set(echo = FALSE)
```


## Generating Sequences Part 1


### Lesson objectives
- Generating sequences using R code

### Packages used in this lesson:

- `tidyverse`

### Success criteria
* Generate different sequences using R code

### Keywords
* sequence
* term = the numbers in the sequence
* nth term

### Worked Example 1

First we will tell R to generate a sequence.

This is how you generate a sequence of numbers from 1 to 100 where the terms increase by 1.
```{r chunk11, exercise=TRUE}
seq(from=1,to=100,by=1)
```

Use the above example R code in chunk1 to generate the following sequences.  Remember to edit the numbers in the code. 

#### Activity 1 Questions
Q1 Generate a sequence from 1 to 100 where the terms increase by 10

```{r chunk21, exercise=TRUE}
```

Q2 Generate a sequence from 5 to 95 where the terms increase by 5.
```{r chunk31,  exercise=TRUE}
```

Q3 Generate a sequence from 3 to 399 where the terms increase by 4.
```{r chunk41,  exercise=TRUE}
```

Q4 Generate a sequence from 1 to 9 where the terms increase by 3.
```{r chunk51,  exercise=TRUE}
```


### Worked Example 2
Now we will generate a sequence that starts from a number and goes up to the nth term, where the nth term is for example 10.

For example a sequence from 1 where the terms go up by 1 till the 10th term.
```{r chunk61,  exercise=TRUE}
seq(from=1, length.out=10, by=1)
```

Use the above example R code in chunk6 to generate the following sequences.  

Knit your document after each question to check your sequence.

#### Activity 2 Questions
Q1 Generate a sequence from 1 where the terms go up by 4 up to the 100th term

```{r chunk71,  exercise=TRUE}
```

Q2 Generate a sequence from 5 where the terms go up by 10 up to the 89th term
```{r chunk81,  exercise=TRUE}
```

Q3 Generate a sequence from 3 where the terms go up by 3 up to the 33rd term
```{r chunk91, exercise=TRUE}
```


### Worked Example 3
So the terms have increased by a particular number, but we can also decrease by a number and produce a backward sequence.

For example, we can sequence from 100 to 1 by -1

```{r chunk101, exercise=TRUE}
seq(from=100, to=1, by=-1)
```


Use the above example R code in chunk10 to answer the following questions.

Knit your document after each question to check your sequence is correct.

#### Activity 3 Questions

Q1 Generate a sequence that goes down from 50 to 0 where the terms go down by 5.
```{r chunk111, exercise=TRUE}
```

Q2 Generate a sequence from 10 to -10 where the terms go down by 1.
```{r chunk121, exercise=TRUE}
```

Q3 Generate a sequence that goes down from 20 to 10 where the terms go down by 2.
```{r chunk141, exercise=TRUE}
```


## Generating Sequences Part 2

### Lesson objectives
- Investigating sequences

### Packages used in this lesson:

- `tidyverse`

### Success criteria
* Find the relationship between the terms
* Does a number lie in a sequence
* Find a missing term in the sequence

### Keywords
* sequence
* term
* nth term


#### Worked Example 1
This is how you find the difference between the terms in a sequence.

The sequence is 3, 7, 11, 15.

Knit your document to see the sequence generated by the R code in chunk1.

```{r chunk1, exercise=TRUE}
x <- c(3,7,11,15)
diff(x)
```

The R code tells you the difference between each number in the sequence.

##### Activity 1 Questions

Now use the above example R code in chunk1 to find the difference between the terms in the following sequences

Knit the document after each question to see the answer.  Check your answer.

Q1  2,7,12,17,22
```{r chunk2, exercise=TRUE}
```

Q2  3,9,15,21,27
```{r chunk3, exercise=TRUE}
```

Q3 1,4,10,19,31
```{r chunk4, exercise=TRUE}
```


#### Worked Example 2
Now we will see if a number lies in a sequence.

First we will generate the first 20 terms of the sequence 3,7,11,15 and then we will see if the numbers 55 and 56 lie in the sequence. The computer will generate the sequence and if it finds the number it will report it below the sequence.  If it does not find it then it will say numeric (0).

```{r chunk5, exercise=TRUE}
x <- seq(from=3, length.out=20, by=4)
x
x[x==55]
x[x==56]
```

Use the above example R code in chunk5 to answer the following questions.

Remember to tell R how many terms you want it to generate by editing the length.out number. Remember to tell R what numbers to look for in the sequence by editing the x== numbers.


##### Activity 2 Questions
Q1 Generate the first 100 terms of the sequence 3,7,11,15 and see if the numbers 117 and 395 lie in the sequence. Do they?

```{r chunk6, exercise=TRUE}
```


#### Worked Example 3
Now we will find the missing number in a sequence.  Our sequence is 4, ?, 16, 22. Term 2 is missing.  So we generate the sequence based on what we know which is that the difference between 16 and 22 is 6.  The computer will put in the missing term.  

```{r chunk7, exercise=TRUE}
seq(from=4, to=22, by=6)
```
Answer: The missing term is 10.

Use the above example R code to answer the following questions.

Remember to knit to see your results.

##### Activity 3 Questions
Q1 What is the missing term in the sequence 3,12,?,30.

```{r chunk8, exercise=TRUE}
```


## Topic 1

### Exercise 

*Here's a simple exercise with an empty code chunk provided for entering the answer.*

Write the R code required to add two plus two:

```{r two-plus-two, exercise=TRUE}

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

